---
title: "NY county map"
author: "Linnea Powell"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## load packages
library (geojsonio)
library( maptools )
library( sp )
library (dplyr)
library(stringr)
library(ggplot2)
library(censusapi)
library(dplyr)
library(shiny)

```

## 1. Load Data from 3 sources


### Load Counties Map, convert to dataframe, and prep for merge

#### Load map file
```{r}
# Read in parcel geojson file from Github
ny.counties <- geojson_read("https://raw.githubusercontent.com/DataCapstone/Data-Capstone/master/Raw-Data/NY_counties.geojson", method="local", what="sp" )

```

#### Clean geoid to match county code
```{r}
# Restructure geoid to match county code
colnames(ny.counties@data) <- c("geoid","county_name")

ny.counties@data$county_code <- substr(ny.counties@data$geoid, 10, 12)

ny.counties@data$county_code <- as.numeric(as.character(ny.counties@data$county_code))
```



### Load USA Spending Grants Data for New York FY 2016

#### Load Raw Data
```{r}
# Read in raw data of 2016 grants in New York state with rds file
rds_url <- "https://github.com/DataCapstone/Data-Capstone/blob/master/Raw-Data/NYgra16.rds?raw=true"

ny.grants.dat <- readRDS( gzcon(url( rds_url ) ) )
```

#### Exclude Funding to State Government
```{r}
# Exclude funding going to state government
ny.nonstategrants.dat <- subset (ny.grants.dat, recipient_type != "00: State government")
```

#### Select columns
```{r}
# Select columns for recipient county code, and recipient county name, and federal funding amount
counties.fedfunding.dat<- select(ny.nonstategrants.dat, recipient_county_code,recipient_county_name, fed_funding_amount)
```

### Load Population Data from Census API
```{r}
# download acs5 2015 for Population by County
censuskey <- "b06146bd7469912826a96e117fcf4a8ab8bc36de"
population.dat <- getCensus(name="acs5", vintage=2015, key=censuskey, 
                       vars=c("B01001_001E"), region="county:*", regionin = "state: 36")

colnames(population.dat) <- c("state","county_code", "population")

population.dat$county_code <- as.numeric(as.character(population.dat$county_code))
```

## 2. Aggregate by county, merge data, and create per capita funding variable

### Add up all federal funding by county
```{r}
# aggregate funding by county code
grantdollarsbycounty.agg<-aggregate(counties.fedfunding.dat$fed_funding_amount, by=list(counties.fedfunding.dat$recipient_county_code), FUN=sum, na.rm=TRUE)

# rename columns
colnames(grantdollarsbycounty.agg) <- c("county_code","federal_funding")

grantdollarsbycounty.agg$county_code <- as.numeric(as.character(grantdollarsbycounty.agg$county_code))
```

### Merge data for federal funding and population 
```{r}
# merge county population data with aggregated county federal funding data by county code

counties.merged.dat <- merge(population.dat, grantdollarsbycounty.agg, by = "county_code", all.x=TRUE, na.rm=TRUE)
```

### Create new variable for per capita funding
```{r}
counties.merged.percapita.dat<-mutate(counties.merged.dat, funding_per_capita=federal_funding/population)
```

### Select columns needed for plot
```{r}
counties.selected<- select(counties.merged.percapita.dat, county_code, federal_funding, funding_per_capita)
```

### Merge with counties shape file
```{r}

# merge map dataframe with funding data frame by county code to add in geoid and county name

counties.plot<- merge(ny.counties, counties.selected, by= "county_code", all.x=TRUE, na.rm=TRUE)
```


### Save off new rds file
```{r}
# Generate an rds file 
saveRDS (counties.plot, "Counties.plot.funding.aggregated.rds" )
```


### Create data frames ordered by federal funding amount and per capita funding

```{r}
# order by federal funding amount
counties.plot<- counties.plot[order(counties.plot$federal_funding), ]

```


## 3. Plot counties color coded by amount of funding

### Organizing colors
```{r}
# Create a color palette to display amount of federal funding on a map
colors<- colorRampPalette( c("gray100","purple4"))

# Generate colors on that continuum
col.ramp <- colors(31)

# Break up the values of the vectors into levels and cast as character

counties.plot$color <-
cut(rank(counties.plot$federal_funding), breaks = 31, labels=col.ramp)

counties.plot$color<- as.character(counties.plot$color)

```


### Plot County Map of Total Federal Grant Funding

```{r}
plot(counties.plot, main="New York's Federal Grant Funding Totals by County, FY 2016",
    col=counties.plot$color,
    border=0    )
```


### Plot County Map of Per Capita Grant Funding

```{r}
# order by per capita funding amount
counties.plot<- counties.plot[order(counties.plot$funding_per_capita), ]

counties.plot$color2<-
cut(rank(counties.plot$funding_per_capita), breaks = 31, labels=col.ramp)

counties.plot$color2<- as.character(counties.plot$color2)

plot(counties.plot, main="New York's Federal Grant Funding Per Capita by County, FY 2016",
    col=counties.plot$color2,
    border="gray80"    )
```



