---
title: "Contracts dataframe"
output: github_document
html_document:
  df_print: paged
  keep_md: true
  code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F, fig.width = 10, fig.height = 10 )
library(dplyr)
library(pander)
setwd("C:/Users/icps/Dropbox/3. Maxwell/4. Capston/Data-Capstone/ignacio")
cnt <- read.csv("./Data/NYcon2017.csv", stringsAsFactors = F)

cnt$transaction_status

#making naics code two digit
cnt$NAICS.code <- as.integer(substring(cnt$principalnaicscode, 1,2))

#adding the name variable
naics <- read.csv("NAICS_cat.csv")
x <- match(cnt$NAICS.code, naics$X2017.NAICS.Code)
cnt$NAICS.name <- as.character(naics$X2017.NAICS.Title[x])

```


***

#Working by a NAICS code

Interesting NAICS:

* Manufacturing
* Professional, Scientific, and Technical Services
* Health Care and Social Assistance
* Construction
* Educational Services

```{r}
#subsetting Construction NAICS
x <- cnt$NAICS.name == "Construction"
con <- cnt[x, c("piid", "maj_agency_cat", "dollarsobligated", 
                      "contractactiontype", "productorservicecode", "psc_cat",
                      "pop_state_code", "placeofperformancecity", 
                      "competitiveprocedures", "multiyearcontract", "multipleorsingleawardidc", 
                      "vendorname", "NAICS.name", "receivescontracts", "receivesgrants",
                      "receivescontractsandgrants", "annualrevenue", "numberofemployees", "isconstructionfirm")]

#sum of money in Construction
x <- is.na(con$dollarsobligated)
sum(con$dollarsobligated[!x], na.rm = T)

#number of cases
length(con$dollarsobligated[!x])

#top 5 Agencies
arrange(as.data.frame(table(con$maj_agency_cat)), desc(Freq))[c(1:5),]

#Top 5 PSC
arrange(as.data.frame(table(con$productorservicecode)), desc(Freq))[c(1:5),]
arrange(as.data.frame(table(con$psc_cat)), desc(Freq))[c(1:5),] #better to use this code!

#top 5 Vendors
arrange(as.data.frame(table(con$vendorname)), desc(Freq))[c(1:5),]
con.grp <- group_by(con, vendorname)
con.grp <- arrange(summarize(con.grp, Monies = sum(dollarsobligated), Freq = n()), desc(Monies))


#Looking at only the top vendor

x <- cnt$vendorname == "TRIO ASBESTOS REMOVAL CORP." & cnt$NAICS.name == "Construction"
con2 <- cnt[x, c("piid", "maj_agency_cat", "dollarsobligated", 
                      "contractactiontype", "productorservicecode", "psc_cat",
                      "pop_state_code", "placeofperformancecity", 
                      "competitiveprocedures", "multiyearcontract", "multipleorsingleawardidc", 
                      "vendorname", "NAICS.name", "receivescontracts", "receivesgrants",
                      "receivescontractsandgrants", "annualrevenue", "numberofemployees", "isconstructionfirm")]

unique(con2$productorservicecode)


#filtering by isacontructionfirm
unique(as.character(cnt$isconstructionfirm)) #this variable is not useful

```

***

#Working by Agency

Interesting Agencies:

"7300: Small Business Administration" 
"8600: Department of Housing and Urban Development"
"9100: Department of Education"  
"6900: Department of Transportation"

```{r}

```

***

#Working by Vendor

```{r}


```
