---
title: "Grants data frame"
output: github_document
html_document:
  df_print: paged
  keep_md: true
  code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F, fig.width = 10, fig.height = 10 )
library(dplyr)
library(pander)
setwd("C:/Users/icps/Dropbox/3. Maxwell/4. Capston/Data-Capstone/ignacio")
gra16 <- read.csv("./Data/NYgra2016.csv", stringsAsFactors = F)

```


```{r, eval=FALSE}
#generating a csv with the variables
x <- gra16[1,]
x <- as.data.frame(lapply(x, class))
x <- as.data.frame(t(x))
x <- data.frame(var = row.names(x), class = as.character(x$V1), stringsAsFactors = F)
grav <- x
write.csv(grav, "gra.variables.csv", row.names = F)

```

```{r}
#exploring some variables

#types of assistance
#what type of assistances are they getting?
unique(gra16$asst_cat_type)
unique(gra16$assistance_type)

#cfda_program_title
head(gra16$cfda_program_title)
head(gra16$cfda_program_num)
(table(gra16$cfda_program_title == "")) #has 110 blank ones.
table(is.na(gra16$cfda_program_title)) #no NAs here!
class(gra16$cfda_program_title)

#fda_program_num
table(is.na(gra16$cfda_program_num)) #no NAs here!
table(gra16$cfda_program_num == "") #no blanks!
class(gra16$cfda_program_num)

#project_description
unique(gra16$project_description)

######PRINCIPAL PLACE###############
#location: wanna focus on 4 cities, no idea if they are comparable.
sort(unique(gra16$principal_place_cc))
# "Syracuse" | "SYRACUSE"   
# "Rochester" | "ROCHESTER"
# "Albany" | "ALBANY"
#  "Ithaca" | "ITHACA" 

sort(unique(gra16$principal_place_state)) 
#Output is  "Illinois" "Massachusetts" "New York" "NEW YORK"
#this means that the filter in the manual download is not for this variable. 

######RECIPIENTS###############
sort(unique(gra16$recipient_state_code)) 
#output is only NY, so the filter is by recipients place, which makes sense because if we are an IDA we should care about the recipient not the pop.

sort(unique(gra16$recipient_city_name)) 
# "Syracuse" | "SYRACUSE"   
# "Rochester" | "ROCHESTER"
# "Albany" | "ALBANY"
#  "Ithaca" | "ITHACA" 

sort(unique(gra16$recipient_county_code))
sort(unique(gra16$recipient_county_name))
sort(unique(gra16$recipient_state_code))

```



**ASSISSTANCE**

1	unique_transaction_id	character
16	federal_award_id	character

18	fed_funding_amount	numeric
19	non_fed_funding_amount	numeric
20	total_funding_amount	numeric

50	asst_cat_type	character
24	assistance_type	character
25	record_type	integer             :  Federal Assistance Awards Data System record type: 1 = county aggregate record, 2 = individual action record.
47	fiscal_year	integer

23	ending_date	character

28	principal_place_code	character
29	principal_place_state	character
30	principal_place_cc	character
32	principal_place_zip	character
48	principal_place_state_code	character

**AGENCY/PROGRAM**

52	maj_agency_cat	character
35	agency_name	character
4	cfda_program_num	numeric
34	cfda_program_title	character
15	agency_code	character

**RECIPIENT**

36	project_description	character
7	recipient_name	character
13	recipient_type	character
49	recip_cat_type	character

8	recipient_city_code	integer
9	recipient_city_name	character
10	recipient_county_code	integer
11	recipient_county_name	character
56	recipient_state_code	character
12	recipient_zip	integer
42	receip_addr1	character
43	receip_addr2	character
44	receip_addr3	character

57	exec1_fullname	character
58	exec1_amount	numeric

#Questions

1. how to work with negatives?