---
title: "IDA Use Case"
output: github_document
html_document:
  df_print: paged
  keep_md: true
  code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = F, message=F, warning=F, fig.width = 10, fig.height = 10 )
library(dplyr)
library(pander)
library(censusapi)
library(geojsonio)
library(sp) #need this to read the json shapes in gplot

########################

#CENSUS DF

# download acs5 2015 for Population, Median Household income, White, Black, Asian, Hispanic or latino, Unemployed, Income < poverty level last 12 months, In Labor Force
censuskey <- "5a9ad013c3692d7dfdd344a9b2b774f36c50b43d"
dat <- getCensus(name="acs5", vintage=2015, key=censuskey, 
                       vars=c("B01001_001E", "B19013_001E", "B01001A_001E","B01001B_001E", "B01001D_001E", "B01001I_001E", "B23025_005E", "B17001_002E", "B23025_002E"), region="county:*", regionin = "state: 36")

colnames(dat)[3:11] <- c("Pop", "MHincome", "White", "Black", "Asian", "Hispanic", "Unemployed", "Poverty", "Labor")

#creating a county.name variable

#getting a DF with the names of the counties and codes.
cou <- read.csv("https://raw.githubusercontent.com/DataCapstone/Data-Capstone/master/Raw-Data/countycodesNY.csv", stringsAsFactors = F)
dat$county.name <- as.factor(dat$county)
levels(dat$county.name) <- as.character(cou$Name)
#I used levels to standardize the names, but another way to do this could be using the match function:
#order <- match(as.numeric(dat$county), cou$Fips)

#ordering the DF
dat <- dat[,c(1:2,12,3:11)]

########################

#Loading a geojson file for NY counties
shapes <- geojson_read("https://raw.githubusercontent.com/DataCapstone/Data-Capstone/master/Raw-Data/NY_counties.geojson", method="local", what="sp" )
#standardizing the county name in the shapes@data$name
levels(shapes@data$name) <- levels(dat$county.name)

########################
#grants DF

gra16 <- readRDS( gzcon(url("https://github.com/DataCapstone/Data-Capstone/blob/master/Raw-Data/NYgra16.rds?raw=true")))

#gra17 <- read.csv("./Data/NYgra2017.csv", stringsAsFactors = F)
#gra16 <- read.csv("./Data/NYgra2016.csv", stringsAsFactors = F)
#gra15 <- read.csv("./Data/NYgra2015.csv", stringsAsFactors = F)
#gra14 <- read.csv("./Data/NYgra2014.csv", stringsAsFactors = F)

setwd("C:/Users/icps/Dropbox/3. Maxwell/4. Capston/Data-Capstone/ignacio")

```


#Use case

* **Position**: Howard Zemsky president and CEO of Empire State Development
* **Organization**: Empire State Development
* **Location**: NY
* **Sector**: Local Government (State)
* **Interest/question**: Wants to know which cities in their state are missing out on potential funds that they could apply for and needs a breakdown of per capita federal spending in each city to assess areas of need.
 
In what follows I will elaborate comparisons between Rochester, Syracuse, Ithaca, Buffalo and Albany.


```{r}
#Generating subsets for 5 cities 2016 data

#Subsetting sry16
x <- gra16$recipient_city_name %in% c("Syracuse", "SYRACUSE")

dat <- gra16[x,c("maj_agency_cat", "agency_name", "action_type",
               "cfda_program_num", "cfda_program_title", "fed_funding_amount", "non_fed_funding_amount",
               "total_funding_amount",	"asst_cat_type", "assistance_type", "record_type",
               "fiscal_year", "principal_place_cc", "project_description", "recipient_name", 
               "recipient_type", "recip_cat_type", "recipient_city_name", "recipient_county_name",
               "recipient_state_code", "exec1_fullname", "exec1_amount")]

#Only interested in pure Federal grants
x <- dat$non_fed_funding_amount == 0
dat <- dat[x,]

#only interested in project grants
x <- dat$assistance_type == "04: Project grant"
dat <- dat[x,]

#saving it as syr16
syr16 <- dat

###########################

#Subsetting Rochester
x <- gra16$recipient_city_name %in% c("Rochester", "ROCHESTER")

dat <- gra16[x,c("maj_agency_cat", "agency_name", "action_type",
               "cfda_program_num", "cfda_program_title", "fed_funding_amount", "non_fed_funding_amount",
               "total_funding_amount",	"asst_cat_type", "assistance_type", "record_type",
               "fiscal_year", "principal_place_cc", "project_description", "recipient_name", 
               "recipient_type", "recip_cat_type", "recipient_city_name", "recipient_county_name",
               "recipient_state_code", "exec1_fullname", "exec1_amount")]

#Only interested in pure Federal grants
x <- dat$non_fed_funding_amount == 0
dat <- dat[x,]

#only interested in project grants
x <- dat$assistance_type == "04: Project grant"
dat <- dat[x,]

roc16 <- dat
###########################

#Subsetting for Albany
x <- gra16$recipient_city_name %in% c("Albany", "ALBANY")

dat <- gra16[x,c("maj_agency_cat", "agency_name", "action_type",
               "cfda_program_num", "cfda_program_title", "fed_funding_amount", "non_fed_funding_amount",
               "total_funding_amount",	"asst_cat_type", "assistance_type", "record_type",
               "fiscal_year", "principal_place_cc", "project_description", "recipient_name", 
               "recipient_type", "recip_cat_type", "recipient_city_name", "recipient_county_name",
               "recipient_state_code", "exec1_fullname", "exec1_amount")]

#Only interested in pure Federal grants
x <- dat$non_fed_funding_amount == 0
dat <- dat[x,]

#only interested in project grants
x <- dat$assistance_type == "04: Project grant"
dat <- dat[x,]

alb16 <- dat

###########################

#Subsetting for Ithaca
x <- gra16$recipient_city_name %in% c("Ithaca", "ITHACA")

dat <- gra16[x,c("maj_agency_cat", "agency_name", "action_type",
               "cfda_program_num", "cfda_program_title", "fed_funding_amount", "non_fed_funding_amount",
               "total_funding_amount",	"asst_cat_type", "assistance_type", "record_type",
               "fiscal_year", "principal_place_cc", "project_description", "recipient_name", 
               "recipient_type", "recip_cat_type", "recipient_city_name", "recipient_county_name",
               "recipient_state_code", "exec1_fullname", "exec1_amount")]

#Only interested in pure Federal grants
x <- dat$non_fed_funding_amount == 0
dat <- dat[x,]

#only interested in project grants
x <- dat$assistance_type == "04: Project grant"
dat <- dat[x,]

ith16 <- dat


###########################

#Subsetting for Buffalo
x <- gra16$recipient_city_name %in% c("Buffalo", "BUFFALO")

dat <- gra16[x,c("maj_agency_cat", "agency_name", "action_type",
               "cfda_program_num", "cfda_program_title", "fed_funding_amount", "non_fed_funding_amount",
               "total_funding_amount",	"asst_cat_type", "assistance_type", "record_type",
               "fiscal_year", "principal_place_cc", "project_description", "recipient_name", 
               "recipient_type", "recip_cat_type", "recipient_city_name", "recipient_county_name",
               "recipient_state_code", "exec1_fullname", "exec1_amount")]

#Only interested in pure Federal grants
x <- dat$non_fed_funding_amount == 0
dat <- dat[x,]

#only interested in project grants
x <- dat$assistance_type == "04: Project grant"
dat <- dat[x,]

buf16 <- dat

```

#Total Amount of FED money going to the chosen cities

```{r}

x <- data.frame(City = c("Syr", "Roc", "Buf", "Ith", "Alb"), Fed = c(sum(syr16$fed_funding_amount), sum(roc16$fed_funding_amount), sum(buf16$fed_funding_amount), sum(ith16$fed_funding_amount), sum(alb16$fed_funding_amount)))

x<- arrange(x, desc(Fed))

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(x$Fed/1000000, names.arg = x$City, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8, col="#4979FF")
  text(y=x$Fed[5]/1000000+250, x=5, as.character(round(x$Fed[5]/1000000, digits=1)), cex = .7)       

for (i in 1:length(x$Fed)) {
    text(y=x$Fed[i]/1000000+250, x=i, as.character(round(x$Fed[i]/1000000, digits=1)), cex = .7)       
}

  length(unique(gra16$cfda_program_title))


```


#Main Recipient Types (recip_cat_type) of Fed $ in the cities

##Syracuse
```{r}
#Main recipients at syr
options(scipen = 999)
rec <- group_by(syr16, recip_cat_type)
rec <- arrange(as.data.frame(summarize(rec, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#printing head
rec %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(rec$Fed/1000000, names.arg = rec$recip_cat_type, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8, col="#4979FF")

#creating an object for a pie charts for the 5 mayor recipients
pie.syr.rec <- rec

```


#Main Recipients of Fed $ in the cities

##Syracuse
```{r}
#Main recipients at syr
options(scipen = 999)
rec <- group_by(syr16, recipient_name)
rec <- arrange(as.data.frame(summarize(rec, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#removing negatives
x <- rec$Fed < 0 
rec$Fed[x] <- 0
rec <- arrange(rec, desc(Fed))

#printing head
head(rec) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(rec$Fed/1000000, names.arg = rec$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

#creating an object for a pie charts for the 5 mayor recipients
pie <- rec[1:5,]
pie[6,] <- c("OTHERS", sum(rec$Fed[6:length(rec$Fed)]), sum(rec$Freq[6:length(rec$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.syr.rec <- pie

```

##Rochester
```{r}
#Main recipients in ROCHESTER
rec <- group_by(roc16, recipient_name)
rec <- arrange(as.data.frame(summarize(rec, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#removing negatives
x <- rec$Fed < 0 
rec$Fed[x] <- 0
rec <- arrange(rec, desc(Fed))

#printing head
head(rec) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(rec$Fed/1000000, names.arg = rec$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

#creating an object for the pie chart of only 5 mayor recipients
pie <- rec[1:5,]
pie[6,] <- c("OTHERS", sum(rec$Fed[6:length(rec$Fed)]), sum(rec$Freq[6:length(rec$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.roc.rec <- pie
```

##Ithaca
```{r}
#Main recipients in Ithaca
rec <- group_by(ith16, recipient_name)
rec <- arrange(as.data.frame(summarize(rec, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#removing negatives
x <- rec$Fed < 0 
rec$Fed[x] <- 0
rec <- arrange(rec, desc(Fed))

#printing head
head(rec) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(rec$Fed/1000000, names.arg = rec$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

#creating an object for the pie chart of only 5 mayor recipients
pie <- rec[1:5,]
pie[6,] <- c("OTHERS", sum(rec$Fed[6:length(rec$Fed)]), sum(rec$Freq[6:length(rec$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.ith.rec <- pie

```

##Buffalo
```{r}
#Main recipients in BUFFALO
rec <- group_by(buf16, recipient_name)
rec <- arrange(as.data.frame(summarize(rec, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#removing negatives
x <- rec$Fed < 0 
rec$Fed[x] <- 0
rec <- arrange(rec, desc(Fed))

#printing head
head(rec) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(rec$Fed/1000000, names.arg = rec$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

#creating an object for the pie chart of only 5 mayor recipients
pie <- rec[1:5,]
pie[6,] <- c("OTHERS", sum(rec$Fed[6:length(rec$Fed)]), sum(rec$Freq[6:length(rec$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.buf.rec <- pie

```

##Albany
```{r}
#Main recipients in ALBANY
rec <- group_by(alb16, recipient_name)
rec <- arrange(as.data.frame(summarize(rec, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#removing negatives
x <- rec$Fed < 0 
rec$Fed[x] <- 0
rec <- arrange(rec, desc(Fed))

#printing head
head(rec) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(rec$Fed/1000000, names.arg = rec$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

#creating an object for the pie chart of only 5 mayor recipients
pie <- rec[1:5,]
pie[6,] <- c("OTHERS", sum(rec$Fed[6:length(rec$Fed)]), sum(rec$Freq[6:length(rec$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.alb.rec <- pie

```


Some insights:

* Main recipients are universities
* Albany is an outlier because it prob receives funds for the whole state.

##Comparing Top 5 Recipients in Syr and Roc

```{r}
#pie charts!
col <- c("#4979FF", "#FF9858",  "#8DFFFF", "#FF84FF", "#55D473", "#FFFF78")

plot.new()
par(mfrow = c(1,2), mar=c(0,0,3,0))

pie<- pie.syr.rec
pie(as.numeric(pie$Fed), labels = paste0(pie$recipient_name, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 5 Recipients of Fed $ in Syracuse", cex= .7, col=col)

pie<- pie.roc.rec
pie(as.numeric(pie$Fed), labels = paste0(pie$recipient_name, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 5 Recipients of Fed $ in Rochester", cex= .7, col=col)

```

#Main Fed Agencies and Programs that give funding

##Syracuse

###Agencies
```{r}
#Syracuse

#First by Agencies
dat <- group_by(syr16, agency_name)
dat <- arrange(as.data.frame(summarize(dat, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#pie chart only 5 mayor recipients
pie <- dat[1:5,]
pie[6,] <- c("OTHERS", sum(dat$Fed[6:length(dat$Fed)]), sum(dat$Freq[6:length(dat$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.syr.age <- pie

#removing negatives
x <- dat$Fed < 0 
dat$Fed[x] <- 0
dat <- arrange(rec, desc(Fed))

#printing head
head(dat) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(dat$Fed/1000000, names.arg = dat$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

```

###Programs
```{r}
#Main programs
dat <- group_by(roc16, cfda_program_title)
dat <- arrange(as.data.frame(summarize(dat, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

pie <- dat[1:10,]
pie[11,] <- c("OTHERS", sum(dat$Fed[11:length(dat$Fed)]), sum(dat$Freq[11:length(dat$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.syr.pro <- pie

#removing negatives
x <- dat$Fed < 0 
dat$Fed[x] <- 0
dat <- arrange(rec, desc(Fed))

#printing head
head(dat) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(dat$Fed/1000000, names.arg = dat$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

```


##Comparing Top Funding Agencies and Programs in Syr and Roc

```{r}
#plotting both pies

par(mfrow=c(2,1), mar=c(0,0,3,0))
pie <- pie.syr.age
pie(as.numeric(pie$Fed), labels = paste0(pie$agency_name, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 5 Agencies of Fed $ in Syracuse", cex= .7, col=col)

pie <- pie.syr.pro
pie(as.numeric(pie$Fed), labels = paste0(pie$cfda_program_title, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 10 Programs of Fed $ in Syracuse", cex= .7, col=col)

```

##Rochester

###Agencies
```{r}
#ROCHESTER

#First by Agencies
dat <- group_by(roc16, agency_name)
dat <- arrange(as.data.frame(summarize(dat, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#pie chart only 5 mayor recipients
pie <- dat[1:5,]
pie[6,] <- c("OTHERS", sum(dat$Fed[6:length(dat$Fed)]), sum(dat$Freq[6:length(dat$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.roc.age <- pie

#removing negatives
x <- dat$Fed < 0 
dat$Fed[x] <- 0
dat <- arrange(rec, desc(Fed))

#printing head
head(dat) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(dat$Fed/1000000, names.arg = dat$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

############################
```

###Programs
```{r}

#Main programs
dat <- group_by(roc16, cfda_program_title)
dat <- arrange(as.data.frame(summarize(dat, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

pie <- dat[1:10,]
pie[11,] <- c("OTHERS", sum(dat$Fed[11:length(dat$Fed)]), sum(dat$Freq[11:length(dat$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))
pie.roc.pro <- pie

#removing negatives
x <- dat$Fed < 0 
dat$Fed[x] <- 0
dat <- arrange(rec, desc(Fed))

#printing head
head(dat) %>% pander

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,1))
barplot(dat$Fed/1000000, names.arg = dat$recipient_name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8)

############################

#plotting both pies

par(mfrow=c(2,1), mar=c(0,0,3,0))
pie <- pie.roc.age
pie(as.numeric(pie$Fed), labels = paste0(pie$agency_name, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 5 Agencies of Fed $ in Rochester", cex= .7, col=col)

pie <- pie.roc.pro
pie(as.numeric(pie$Fed), labels = paste0(pie$cfda_program_title, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 10 Programs of Fed $ in Rochester", cex= .7, col=col)



```

###Comparable Pie charts for Agencies and Programs in both Syr and Roc

```{r}
#all four 
par(mfrow=c(2,2), mar=c(0,0,3,0))
pie <- pie.syr.age
pie(as.numeric(pie$Fed), labels = paste0(pie$agency_name, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 5 Agencies of Fed $ in Syracuse", cex= .7, col=col)

pie <- pie.syr.pro
pie(as.numeric(pie$Fed), labels = paste0(pie$cfda_program_title, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 10 Programs of Fed $ in Syracuse", cex= .7, col=col)

pie <- pie.roc.age
pie(as.numeric(pie$Fed), labels = paste0(pie$agency_name, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 5 Agencies of Fed $ in Rochester", cex= .7, col=col)

pie <- pie.roc.pro
pie(as.numeric(pie$Fed), labels = paste0(pie$cfda_program_title, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Top 10 Programs of Fed $ in Rochester", cex= .7, col=col)

```


#Total Amount Received as compared to other cities


```{r}
unique(syr16$agency_name)
unique(roc16$agency_name)



```




```{r}
#Using the census data we can visualize the Population of NY state by county. 

par(mar = c(7,6,3,3), mfrow=c(1,1))
dat <- arrange(dat, desc(Pop))
barplot(dat$Pop/1000, names.arg = dat$county.name, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in 1000 USD", cex.lab = .8, col="dodgerblue4")

#Generating a color pallette
colors <- colorRampPalette( c("white", "light grey","dodgerblue4")) #this makes a function called "col"
colors <- colors( 7 ) # using the function we input the arg for the number of colors in that continium. And then saving the output as col too.

#dividing the ranked data into groups and adding color as a label. 
col <- cut(rank(dat$Pop), breaks=7, labels=colors)
col <- as.character(col) #the result is a list corresponding to the county but with the pop color.

#making the label for our legend.
#brakets <- seq(from=1, to = max(dat$Pop), by = max(dat$Pop)/7)
#1.0  370752.3  741503.6 1112254.9 1483006.1 1853757.4 2224508.7
brakets <- c("< 371k", "371k - 741k", "741k - 1.1m", "1.1m - 1.5m", "1.5m - 1.8m", "1.8m - 2.3m")

#plotting the graph
par(mar= c(0,1,1,1))
plot(shapes, main="County Population in New York State",
    axes = F,
    cex.axis=.8,
    col=col 
    )

#plotting the legend
legend("right", bg="white"
        , cex=0.8
        , legend= rev(brakets)
        , fill=rev(colors)
        , box.col="white"
        , title="Population" 
        )

```


 
 
 
 
 
 
 
**Dataset**:
 
**Relevant Variables**:
 
**Description of Analysis**:
 
**Coded Tools needed**:
 
**Visualizations**:
 
**Journey map**:
 
**Final insights/answers to the interest/question/other added value**:
 
**Limitations/further work**:



